---
title: "Final Project- Life Expectancy Estimation Analysis"
author: "Leo Hong"
date: "May 09, 2019"
output: 
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 1.Introduction
 


```{r warning=FALSE, meassage=FALSE, error=FALSE, echo=FALSE}

setwd ('C:/Users/SCG05/Desktop/BigData')

#install tidyverse package

library(tidyverse)


```
## 2.Data analysis on world trade file 



```{r warning=FALSE, meassage=FALSE, error=FALSE, echo=FALSE}

setwd ('C:/Users/SCG05/Desktop/BigData')
#------------------world trade--------------------
#bring file world_trade
world_trade <-read_csv ("world_trade.csv")
#clean (row containing 0 followed by round off e values)
world_trade_new <-filter (world_trade, Imports>0, Exports>0)
options(scipen=999)
#plot 

ggplot(data=world_trade_new, mapping= aes(x= year, y=Imports, color=country))+ geom_line()

ggplot(data=world_trade_new, mapping= aes(x= year, y=Exports, color=country))+geom_line()
#basic regression
world_trade_new1 <-lm (Trade ~ year + Imports+ Exports, data=world_trade_new)
summary(world_trade_new1)
# plot the regression (check the outlier= 292 and residuals)
plot(world_trade_new1)



```

## 3.Data analysis on world GDP file


```{r warning=FALSE, meassage=FALSE, error=FALSE, echo=FALSE}

setwd ('C:/Users/SCG05/Desktop/BigData')
#---------------------world_gdp--------------------
#bring file world_gdp
world_gdp <-read_csv ("world_gdp.csv")
#clean (row containing 0 followed by round off e values)
world_gdp_new <- filter(world_gdp,Unemployment>0)
options(scipen=999)
#plot 

ggplot(data=world_gdp_new, mapping= aes(x= year, y=Unemployment, color=country))+ geom_line()

ggplot(data=world_gdp_new, mapping= aes(x= year, y=GDP_capita, color=country))+ geom_line()
#basic regression
world_gdp_new1 <-lm (Unemployment ~ year + GDP+ GDP_capita, data=world_gdp_new)
summary(world_gdp_new1)
#plot the regression
plot(world_gdp_new1)


```

## 4.Data analysis on world population file



```{r warning=FALSE, meassage=FALSE, error=FALSE, echo=FALSE}

setwd ('C:/Users/SCG05/Desktop/BigData')
#-------------------- world population-----------------------
#bring file world_population
world_population <- read.csv ("world_population.csv")
#clean (row containing 0 followed by round off e values)
world_population_new <-filter(world_population, Mortality>0, Fertility>0)
options(scipen=999)
#plot

ggplot(data=world_population_new, mapping= aes(y=Life_exp, x=year, color=country))+geom_line()

ggplot(data=world_population_new, mapping= aes(y=Mortality, x=year, color=country))+geom_line()

#population may not contribute to Life_exp, so delete population)
world_population_new1 <-select(world_population_new, -Population) %>% distinct
#basic regression
world_population_regression <-lm(Life_exp ~ year+ Mortality + Fertility, data=world_population_new1)
summary(world_population_regression)
#plot the regression
plot(world_population_new1)


```


## 5. Merging all three tables


```{r warning=FALSE, meassage=FALSE, error=FALSE, echo=FALSE}

setwd ('C:/Users/SCG05/Desktop/BigData')
#-------------------------------merge tables-----------------------
#delete country and year from world_gdp first
world_gdp_select <-select (world_gdp, - country, -year)

#left join with world population
world_total <-left_join(world_population, world_gdp)
view(world_total)

#left join with world_trade
world_final <- left_join (world_total, world_trade)
view(world_final)

#delete all rows with 0
world_final <- filter(world_final,Unemployment>0, Fertility>0, Mortality>0, Life_exp>0)

#get rid of population
world_final1 <-select(world_final, -Population)


```

## 6. Run regression on merged file 


```{r warning=FALSE, meassage=FALSE, error=FALSE, echo=FALSE}

setwd ('C:/Users/SCG05/Desktop/BigData')

# run regression (world)

world_final_regression <- lm (Life_exp ~ year+Mortality +Fertility+GDP+GDP_capita+Unemployment +Imports+Exports+Trade, data=world_final)
summary(world_final_regression)
plot (world_final_regression)

```

## 7. Remove outliers


```{r warning=FALSE, meassage=FALSE, error=FALSE, echo=FALSE}

# found outliers (#145, 147, 148)
#get rid of outliers 
world_final1[145,]
world_final1[147,]
world_final1[148,]
world_final1 <- world_final1[-145,, drop=FALSE]
world_final1 <- world_final1[-147,, drop=FALSE]
world_final1 <- world_final1[-148,, drop=FALSE]

```

## 8. Run regression after removing outliers



```{r warning=FALSE, meassage=FALSE, error=FALSE, echo=FALSE}

world_final_regression_new <- lm (Life_exp ~ year+Mortality +Fertility+GDP+GDP_capita+Unemployment +Imports+Exports+Trade, data=world_final1)
summary(world_final_regression_new)
plot (world_final_regression_new)

```

## 9. Best subset regression


```{r warning=FALSE, meassage=FALSE, error=FALSE, echo=FALSE}

setwd ('C:/Users/SCG05/Desktop/BigData')
library(tidyverse)
library(leaps)
#best subset regression using all variables, pick only 4 best)
best <- regsubsets(Life_exp ~ year+Mortality +Fertility+GDP+GDP_capita+Unemployment +Imports+Exports+Trade, data=world_final1)
summary(world_final1, nvmax=9)
summary(best)

#Pick the lowest BIC value --> 6 variables are the best
res.sum <- summary(best)
data.frame(
  Adj.R2 = which.max(res.sum$adjr2),
  CP = which.min(res.sum$cp),
  BIC = which.min(res.sum$bic)
)


```

## 10. Best subset regression analysis



```{r warning=FALSE, meassage=FALSE, error=FALSE, echo=FALSE}
best_subset <- lm(Life_exp ~ Mortality+GDP+GDP_capita+Unemployment+Exports+Trade, data=world_final1)
summary(best_subset)


```

## 11. Interaction Effects



```{r warning=FALSE, meassage=FALSE, error=FALSE, echo=FALSE}

world_final_int <- lm (Life_exp ~ year+Mortality +Fertility+GDP+GDP_capita+Unemployment +Imports+Exports+Trade +(Mortality:Fertility)+ (Imports:Exports)+(GDP:GDP_capita), data=world_final1)
summary(world_final_int)
plot(world_final_int)




```

## 12. Which model is better? (ANOVA test: best subset vs interaction)


```{r warning=FALSE, meassage=FALSE, error=FALSE, echo=FALSE}

anova(best_subset,world_final_int)





```

 